---
title: "Function Practice"
author: "Deanna Elliott"
date: '2023-02-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)
knitr::opts_chunk$set(error=TRUE)
library(tidyverse)
```

One of the equations used to compute automobile fuel efficiency is as follows this is the power required to keep a car moving at a given speed is as follows:

Pb=crolling∗m∗g∗V+12∗A∗pair∗cdrag∗V^3

where crolling and cdrag are rolling and aerodynamic resistive coefficients, typical values are 0.015 and 0.3, respectively.

V: is vehicle speed (assuming no headwind) in m/s (or mps) 

m: is vehicle mass in kg

A is surface area of car (m2)

g: is acceleration due to gravity (9.8m/s2)

p_air = density of air (1.2kg/m3)

Pb is power in Watts

### Write a function to compute power, given a truck of m=31752 kg (parameters for a heavy truck) for a range of different highway speeds.

The surface area of this truck is 16 m2
30 m/s
80 km/hr
120 km/hr
40 km/hr
5 km/hr

```{r}

fuel_eff(A = 16, c_drag = 0.3, c_roll = 0.015, g = 9.8, m = 31752, p_air = 1.2, V = c(30,80, 120, 40, 5))

#217786.32 
#1847963.52 
#5536745.28  
#371021.76   
#23697.72


```


### Plot power as a function of speed

```{r}
power <- c(217786.32, 1847963.52, 5536745.28,  371021.76,   23697.72)
df_power <- data.frame(power)

ggplot(df_power, aes(power)) +
  stat_function(fun = fuel_eff(
```


### Recompute for a lighter vehicle (15,000 kg) Add a second line to your graph for lighter vehicle

```{r}

#recalculate for lighter vehicle

fuel_eff(A = 16, c_drag = 0.3, c_roll = 0.015, g = 9.8, m = 15000, p_air = 1.2, V = c(30,80, 120, 40, 5))
```


Keep your function in its own file, store it in a subdirectory called R and then create a separate Rmarkdown document to use the function and store the Rmarkdown in a different directory