---
title: "Assignment 3 - Looping"
author: "Deanna Elliott"
date: '2023-03-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)

library(tidyverse)
library(purrr)
library(ggpubr)
```

Add an additional super light car with mass 5,000 kg and a surface area 10m2 to possible_cars from looping.Rmd

* car A:  mass 20000 kg, area 22 m^2

* car B: mass 65000 kg, area 30 m^2

* car C: mass 38000 kg, area 22 m^2

*car D: mass 5000 kg, area 10 m^2

```{r create df}

# create data frame for possible cars
possible_cars = data.frame(name = c("A","B","C", "D"), mass=c(10000,65000,38000,5000), area = c(22,30,22,10))

```

Compute the mean and maximum power for Cars A, B, C and for your new car, assuming that average highway speed is 80 km/hr with a standard deviation of 10km/hr; Use 2 different methods to compute the mean and maximum power for each car type; a) FOR loop and b) the pmap function from purr

```{r for loop}

possible_cars

# assume a log normal distribution of speeds with mean 100km/hr
# convert km/hr to m/s
# sample size = 200

nsample = 200
mean_speed = log(80*0.277)

speeds = rlnorm(mean=mean_speed, sd=0.125*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")

# use for loop to cycle through all speeds
# create a data frame to store for loop results: one column for each car and one row for each speed

powerall1 = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))
# because we don't initialize it - values are NA
head(powerall1)

# index in to a matrix (powerall1)
for (i in 1:nrow(possible_cars)) {
  powerall1[,i] = autopower(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}

head(powerall1)
# we can add column names
colnames(powerall1)=possible_cars$name


# convert columns to rows for plotting
powerallr1 = powerall1 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr1)

mean(powerall1$A)    # 112,902 W
mean(powerall1$B)    # 329,503 W
mean(powerall1$C)   # 208,598 W
mean(powerall1$D)   # 52,873 W

max(powerall1$A)    # 858,075 W
max(powerall1$B)    # 1,608,060 W
max(powerall1$C)    # 1,096,820 W
max(powerall1$D)    # 393,910 W

```

```{r pmap}

# use same speed distribution as above
powerall2 = pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds)

# make data frame and add column names
powerall2 = as.data.frame(powerall2, col.names=possible_cars$name)

# use mapply to determine power
powerall2b = mapply(FUN=autopower, A = possible_cars$area, m=possible_cars$mass, MoreArgs = list(V=speeds))
# add column names
colnames(powerall2b)=possible_cars$name

# for plotting sometimes its useful to turn columns in to rows
powerallr2 = powerall2 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr2)

mean(powerall2$A)    # 112,902 W
mean(powerall2$B)    # 329,503 W
mean(powerall2$C)   # 208,598 W
mean(powerall2$D)   # 52,873 W

max(powerall2$A)    # 858,075 W
max(powerall2$B)    # 1,608,060 W
max(powerall2$C)    # 1,096,820 W
max(powerall2$D)    # 393,910 W

```

Create two boxplots (one for each method (FOR and pmap)) that show the range of power consumption (across sampled speeds) for each car type.

```{r plotting}

for_loop_plot = ggplot(powerallr1, aes(car,power, fill=car)) + 
  geom_boxplot() +
  ggtitle("Car Power from For Loop")

mapply_plot = ggplot(powerallr2, aes(car,power, fill=car)) + 
  geom_boxplot() + 
  ggtitle("Car Power from pmap")

# put plots side by side for comparison
ggarrange(for_loop_plot, mapply_plot)
```

