---
title: "2/27 Notes"
author: "Deanna Elliott"
date: '2023-02-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r speed example loop}

speeds = runif(min=0, max=100, n=300)

maxspeed=0
for ( i  in 1:length(speeds)) {
  maxspeed = ifelse(speeds[i] > maxspeed, speeds[i], maxspeed)
}

maxspeed
```
```{r}

max(speeds)
```
FERTILIZER EXAMPLE

Try this - make a random sample of fertilizer application, 10 values with mean of 5 and standard deviation of 0.5

Task 1: Use a for loop to compute the total yield after 10 years
TP is 20cm

```{r 10 years}

fertilizer <- rnorm(n = 10, mean = 5, sd = 0.5)
TP = 20

crop = data.frame(fertilizer=fertilizer)
crop$yield = NA

for (yr in 1:10) {
  crop$yield[yr] = 1.8*fertilizer[yr]^2 - 0.5*fertilizer[yr]  + 0.1*TP
}

sum(crop$yield)

```


Task 2: create a function to that takes as input a single value for annual fertilizer and annual total precipitation and returns yield

```{r annual yield}

# calc_yield = function(fertilizer, TP){
#  yield = (1.8*fertilizer^2)-0.5*fertilizer + 0.1*TP
#  return(yield)
# }

calc_yield(fertilizer, TP)

sum(calc_yield(fertilizer, TP))
```


Task 3: modify your function so that it returns total yield if the user inputs a vector of fertilizer and a vector of precipitation

Check that function returns the same result as your original for loop

Add some error checking and test


```{r vector yield}

# calc_total_yield = function(fertilizer, TP){
#  yield = (1.8*fertilizer^2)-0.5*fertilizer + 0.1*TP
#  total = sum(yield)
#  return(total)
# }

calc_total_yield(fertilizer, TP)

```

