---
title: "3/6 Notes"
author: "Deanna Elliott"
date: '2023-03-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r rare fish}

possible.fish = c("salmon","steelhead","shark","tuna","cod")

catch1 = base::sample(possible.fish, size=20, replace=T)

rarestfish = names(which.min(summary(as.factor(catch1))))
rarestfish
```

```{r graph rare fish}

plottitle = sprintf("The rarest fish is %s", rarestfish)

ggplot(data.frame(catch1=catch1), aes(catch1, fill=catch1))+geom_histogram(stat="count")+labs(title=plottitle)
```

## Flow Control

### Correlation Test Example

```{r}

mycortest  = function(x,y, thresh=0.8) {
  # compute correlation
  res = cor(x,y)
  classification = ifelse(res > thresh, "GOOD", "NotGood")
  return(classification)
}

a = runif(min=1, max=100, n=100)
b = runif(min=1, max=100, n=100)

mycortest(a,b)
```

```{r}

mycortest(a,a)
```

```{r}

# this doesn't work - why??  Unequal lengths for x and y 
mycortest(a, 1)
```

```{r}

# useful to add error checking
mycortest  = function(x,y, thresh=0.8) {
  # compute correlation
  if(length(x) != length(y)) {stop("unequal lengths for x and y")}
  res = cor(x,y)
  classification = ifelse(res > thresh, "GOOD", "NotGood")
  return(classification)
}
mycortest(a, 1)
```


### With A Simple If

```{r}

# Simple "IF*
# imagine we are trying to get a tuna  - lets "fish" by sampling
possible.fish = c("salmon","steelhead","shark","tuna","cod")
catch1 = base::sample(possible.fish, size=1, replace=T)
catch1

ifelse(catch1 == "tuna", "success", "tryagain")

catch1 = "tuna"
ifelse(catch1 == "tuna", "success", "tryagain")
```

## Multiple Alternatives

```{r}

fish = "steelhead"

case_when((fish =="steelhead")  ~ "A", (fish =="tuna") ~ "B", !(fish == "tuna") & !(fish=="steelhead") ~ "C")
```

```{r}

# apply to all of our ocean
# start by making a function
classify_fish = function(fish) {
 class=  case_when((fish =="steelhead")  ~ "A", (fish =="tuna") ~ "B", !(fish == "tuna") & !(fish=="steelhead") ~ "C")
 return(class) }

classify_fish("steelhead")
```

```{r}

classify_fish("cod")
```

```{r}

# apply to our ocean
possible.fish = c("salmon","steelhead","shark","tuna","cod")
catch2 = base::sample(possible.fish, size=20, replace=T)
catch2

classify_fish(catch2)
```

## Design A Function

Inputs:
- Nutrient concentration for at least 5 days
- Temperature for at least 5 days
- Thresholds for temperature and nutrient (with default values)

Output:
- Mean Nutrient Concentration
- Pollution Risk as “low”, “med” or “high”

Take a look at classify_lake.R

example of while
example of flow control with if..else and case_when

```{r}

source("C:/Users/deann/Box/Winter 2023/esm_262/R/classify_lake.R")

# generate some data to try the function
temperature = runif(min=4, max=35, n=30)
nutrient = runif(min=5, max=40, n=30)

# try it
res=classify_lake(temperature=temperature, nutrient=nutrient)

# lets pick some numbers we know should give us a low value
nutrient = runif(min=0, max=5, n=30)
classify_lake(temperature=temperature, nutrient=nutrient)
```

```{r}

# now try high
# use repeat to generate the same value multiple times
nutrient = rep(50, times=30)
temperature = rep(25, times=30)
classify_lake(temperature=temperature, nutrient=nutrient)
```






